---
title: "stat_inf_project_2"
author: "Jerry Dumblauskas"
date: "April 10, 2016"
output: html_document
---

Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
data(ToothGrowth)
```
Provide a basic summary of the data.
```{r}

str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)
unique(ToothGrowth$supp)
unique(ToothGrowth$dose)
boxplot(len~supp*dose, data=ToothGrowth, xlab="Vitamin C Dose", col=(c("gold","darkgreen")))
```

looking at the data (per dose) it is close, so let's say "different supp types have no effect on length" (H0)
```{r}
t.test(len ~ supp, data = ToothGrowth)

```

Here, the p is 6 percent (kind of low >23 percent chance of messing up), but the confidence interval contains 0, so we cannot make a clear acception or rejection of H0

Dose has an impact on tooth growth (.5 v 2)
```{r}
d_05 <- ToothGrowth[which(ToothGrowth$dose==.5),1]
d_20 <- ToothGrowth[which(ToothGrowth$dose==2),1]
t.test(d_05, d_20, paired=FALSE, var.equal=FALSE)
```
low p, test contains 0 -- so this H0 is accepted!

#Conclusions
the higher the dose, the more growth.  The basic test showed no statistically different data between the supp types.  In future analysis, if the dose type is restricted (like getting rid of low dose), it may be possible to infer differences in the supplement types.